---

      

---

```{r, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(mapproj)
library(ggmap)
#library(reshape)
#library(vcd)
library(plotluck)
library(PerformanceAnalytics)
library(psych)
library(scales)
library(car)
library(labelled)

brfss = read.csv("C:/Users/Biljana/Desktop/brfss_final_load_R.csv")
```




# What is the BRFSS?

##
- Nation's premier system of health survey measuring 
    - health related risk behaviors,
  
    - chronic conditions
  
    - preventative measures
  
- Survey Methodology:
    - Telephone Survey 
    
    - Prospective observational study 
  
    - Stratified sampling method
  
    - Random sampling, not random assignment 


#Analysis Overview

##

- Used one year dataset, not prevalence data 
    - 500k Observations and 330 variables 
    - 80 Relevant variables were extracted
    - Focus on Respondents' Perception of Health and what factors impact it  

- Analysis structure 
    - Socio- Economic factors 
    - Obesity and exercise
    - Chronic diseases 
    - Behavioral Risks 

#Socio- Demographic Characteristics

##
```{r, echo = FALSE}
brfss1 = brfss %>% 
  dplyr::select(X_state,Region,marital,children,educa,employ1,income2,renthom1,sex,X_imprace,
X_age_g,htin4,htm4,X_chldcnt,X_educag,X_incomg,X_age80,exerany2,X_rfbmi5,actin11_,pa1min_,exract11,drnkany5,X_rfdrhv4,addepev2,daily_veg,fruit_cons_category,sleptim1,X_bmi5cat,genhlth,X_rfhlth,genhlth ) 

brfss1= brfss1[complete.cases(brfss1), ]

ggplot(brfss1,aes(x=as.factor(X_state),fill=sex)) + 
  geom_bar(data=subset(brfss1,sex=="Female")) + 
  geom_bar(data=subset(brfss1,sex=="Male"),aes(y=..count..*(-1))) + 
  scale_y_continuous(breaks=seq(-60000,90000,10000),labels=abs(seq(-60000,90000,10000))) + 
  coord_flip()+labs(x="State",y="Number of responders")+annotate("text", x ="Texas" , y = 1000, label = "We are here!",fontface="italic",size=3)+theme_classic()+theme(axis.text.x = element_text(angle = 45, hjust = 1))+ggtitle("Where do our responders live?")

```

##
```{r, echo = FALSE}
ggplot(brfss1,aes(x=as.factor(X_age_g),fill=sex)) + 
  geom_bar(data=subset(brfss1,sex=="Female")) + 
  geom_bar(data=subset(brfss1,sex=="Male"),aes(y=..count..*(-1))) + 
  scale_y_continuous(breaks=seq(-60000,90000,10000),labels=abs(seq(-60000,90000,10000))) + 
  coord_flip()+labs(x="Age",y="Population")+ annotate("text", x ="Age 65 or older" , y = 30000, label = "This is my grandma!",fontface="italic")+theme_classic()+theme(axis.text.x = element_text(angle = 45, hjust = 1))+ggtitle("Most of our responders are senior citizens")
```

##

```{r, echo= FALSE}
brfss1$X_educag_factor = factor(brfss1$X_educag, levels = c("Did not graduate high school", "Graduated high school","Graduated from college or technical school","Attended college or technical school"))

total_grp = brfss1 %>% 
              group_by(X_educag_factor) %>%
              count(X_educag_factor)
perc = brfss1%>% 
         group_by(Region, X_educag_factor)  %>% 
         count(Region)
perc = merge(perc, total_grp, by = "X_educag_factor")
perc = mutate(perc, Percentage = n.x/n.y)
ggplot(perc, aes(x = Region, y = Percentage, fill = X_educag_factor)) + 
  geom_bar(stat= "identity", position = "dodge") + 
  scale_y_continuous(labels = percent)+
  facet_grid(.~X_educag_factor,labeller = label_wrap_gen(width=15)) + 
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3.3)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none"
        ) +ggtitle("40% of responders who did not graduate from high school live in the South")
```

##
```{r, echo = FALSE}
brfss1$income2_factor = factor(brfss1$income2, levels = c("Less than $10,000", "Less than $15,000", "Less than $20,000", "Less than $25,000", "Less than $35,000","Less than $50,000","Less than $75,000","$75,000 or more"))

total_grp = brfss1 %>%
              group_by(income2_factor) %>%
              count(income2_factor)

perc = brfss1 %>%
         group_by(Region, income2_factor)  %>%
         count(Region)
perc = merge(perc, total_grp, by = "income2_factor")
perc = mutate(perc, Percentage = n.x/n.y)

ggplot(perc, aes(x = Region, y = Percentage,fill = income2_factor)) +
  geom_bar(stat= "identity", position = "dodge") +
  scale_y_continuous(labels = percent)+
  facet_grid(.~income2_factor,labeller = label_wrap_gen(width=15)) +
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3.3)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none"
        )+ggtitle("South also has highest percentage of responders below povery line")
```

##
```{r, echo = FALSE}
brfss_reorder = brfss1 %>% dplyr::select(genhlth,marital, X_educag,sex,X_chldcnt, X_incomg,X_age_g, X_imprace, employ1,Region  ) %>%
   mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))

total_grp = brfss_reorder %>% 
              group_by(genhlth) %>%
              count(genhlth)
perc = brfss_reorder %>% 
         group_by(Region, genhlth)  %>% 
         count(Region)
perc = merge(perc, total_grp, by = "genhlth")
perc = mutate(perc, Percentage = n.x/n.y)
ggplot(perc, aes(x = Region, y = Percentage, fill = genhlth)) + 
  geom_bar(stat= "identity", position = "dodge") + 
  scale_y_continuous(labels = percent)+
  facet_grid(.~genhlth) + 
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3.3)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none"
        )+ggtitle("South region responders with higher self rated health percentages") 
```


##
```{r, echo = FALSE}
brfss_reorder2 =brfss1 %>% dplyr::select(genhlth, marital, X_educag,sex,X_chldcnt, X_incomg,X_age_g, X_imprace, employ1,Region) %>%
   mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))
total_grp_edu = brfss_reorder2 %>% 
              group_by(genhlth) %>%
              count(genhlth)
perc_edu = brfss_reorder2 %>% 
         group_by(X_educag, genhlth)  %>% 
         count(X_educag)
perc_edu = merge(perc_edu, total_grp_edu, by = "genhlth")
perc_edu = mutate(perc_edu, Percentage = n.x/n.y)
ggplot(perc_edu, aes(x = X_educag, y = Percentage, fill = genhlth)) + 
  geom_bar(stat= "identity", position = "dodge") + 
  scale_y_continuous(labels = percent)+
  scale_x_discrete(labels=c("Attended college or technical school"="College attended","Did not graduate high school"="High school attended","Graduated from college or technical school"="College graduated","Graduated high school"="High school gruadated"))+
  facet_grid(.~genhlth) + 
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
        ) +labs(x="Education")+ggtitle("Does education make you feel healthier?")
```

#Health Perception and Obesity

##
```{r, echo=FALSE}
gHlthOb = brfss %>% 
   filter(!is.na(X_bmi5cat), !is.na(X_rfbmi5),!is.na(X_state), !is.na(Region), !is.na(BMI_in) , !is.na(genhlth), !is.na(poorhlth), !is.na(X_rfhlth),
          !is.na(physhlth),!is.na(menthlth), !is.na(X_age80) , !is.na(sex)) %>% 
    dplyr::select(X_bmi5cat,Obesity = X_rfbmi5, X_state,  BMI_in, Region , genhlth,poorhlth,physhlth, menthlth, X_age80, sex ,Health = X_rfhlth) %>% 
    mutate(X_bmi5cat = ordered(factor(X_bmi5cat, levels = c("Underweight", "Normal weight", "Overweight", "Obese")))) %>% 
    mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))

mosaicplot(~ Health + Obesity, 
           main = "50% of overweight and obese respondents report good health",
           data = gHlthOb, 
           color = TRUE, las = 1, shade = TRUE)
```


##  
```{r, echo=FALSE}
ggplot(gHlthOb, aes(genhlth,BMI_in ))+
    geom_boxplot(aes(fill = genhlth),alpha = .6,size = 1)+
    scale_fill_brewer(palette = "Set2") + 
    geom_point(stat = "summary", fun.y = "mean", color = "red", size = 4)+
    ggtitle("No difference in BMI between Poor and Good Health")+
    labs(x = "Health", y = "BMI", fill = "Health")+
    theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(),
        legend.position = "none"
        ) 
```

## 
```{r, echo=FALSE}
ggplot(gHlthOb, aes(genhlth,poorhlth ))+
    geom_boxplot(aes(fill = genhlth),alpha = .6,size = 1)+
    geom_point(stat = "summary", fun.y = "mean", color = "red", size = 4)+
    ggtitle("Number of days in month overall health was poor better indicator of health")+
    labs(x = "Health", y = "# days poor health")+
    theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(),
        legend.position = "none"
        )
``` 
 
## 
 
```{r, echo=FALSE}
ggplot(gHlthOb, aes(genhlth,physhlth ))+
    geom_boxplot(aes(fill = genhlth),alpha = .6,size = 1)+
    geom_point(stat = "summary", fun.y = "mean", color = "red", size = 4)+
    ggtitle("Days in month overall physical health was poor is a better indicator of health")+
    labs(x = "Health", y = "# days poor physical health")+
    theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(),
        legend.position = "none"
        )
```

## 
```{r, echo=FALSE}
ggplot(gHlthOb, aes(genhlth,menthlth ))+
    geom_boxplot(aes(fill = genhlth),alpha = .6,size = 1)+
    geom_point(stat = "summary", fun.y = "mean", color = "red", size = 4)+
    ggtitle("Number of days in month overall mental health was poor")+
    labs(x = "Health", y = "# days poor mental health")+
    theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(),
        legend.position = "none"
        )
```

##
<img src = "C:/Users/Biljana/Desktop/bmi_funny.jpg" alt = "Image not loaded", width = "768", height = "600" />  

##
```{r, echo = FALSE}
ob1 = brfss %>%
  filter(!is.na(X_rfbmi5),!is.na(physhlth),!is.na(X_pacat1), !is.na(BMI_in), !is.na(X_bmi5cat) ) %>%
  dplyr::select(X_rfbmi5,physhlth ,X_pacat1,BMI_in, X_bmi5cat)

total_grp_ex = ob1 %>% 
              group_by(X_rfbmi5) %>%
              count(X_rfbmi5)

perc_ex = ob1 %>% 
         group_by(X_pacat1, X_rfbmi5)  %>% 
         count(X_pacat1)
perc_ex = merge(perc_ex, total_grp_ex, by = "X_rfbmi5")
perc_ex= mutate(perc_ex, Percentage = n.x/n.y)

ggplot(perc_ex, aes(x = X_pacat1, y = Percentage, fill = X_rfbmi5)) + 
  geom_bar(stat= "identity", position = "dodge") + 
  scale_y_continuous(labels = percent)+
  facet_grid(.~X_rfbmi5) + 
  ggtitle("Obese Respondends report vigorous exercise activity")+
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3)+
  labs(x = "Activity Category", "Percentage")+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
        )

```

##
```{r, echo = FALSE}
extypes = brfss %>%
  filter(!is.na(X_rfbmi5),!is.na(physhlth),!is.na(X_pacat1), !is.na(BMI_in), !is.na(X_bmi5cat),!is.na(exract11),!is.na(pa1min_),!is.na(exerany2), !is.na(actin11_)  ) %>%
  filter(exract11 == "Walking" | exract11 == "Running"|
                    exract11 == "Gardening (spading, weeding, digging, filling)" | 
                    exract11 == "Weight lifting" | exract11 == "Bicycling") %>%
  filter(X_rfbmi5 == "Yes")%>%
  dplyr::select(X_rfbmi5,physhlth ,X_pacat1,BMI_in, X_bmi5cat, exract11, pa1min_, exerany2,actin11_)

total_grp_ex = extypes %>% 
              group_by(X_rfbmi5) %>%
              count(X_rfbmi5)

perc_extype = extypes %>% 
         group_by(exract11, X_rfbmi5)  %>% 
         count(exract11)
perc_extype = merge(perc_extype, total_grp_ex, by = "X_rfbmi5")
perc_extype= mutate(perc_extype, Percentage = n.x/n.y)

ggplot(perc_extype, aes(x = exract11, y = Percentage, fill = X_rfbmi5)) + 
  geom_bar(stat= "identity", position = "dodge") + 
  scale_y_continuous(labels = percent)+
   scale_x_discrete(labels=c("Running" = "Running",
                            "Bicycling" = "Bicycling",
                              "Weight lifting" = "Weight Lifting",
                            "Gardening (spading, weeding, digging, filling)" = "Gardening", 
                            "Walking" = "Walking"))+
  ggtitle("Exercise! Of course I exercise!")+
  labs( x = "exercise Types")+
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), hjust = 0.3, size = 3)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(),
        legend.position = "none"
        )
```

##
```{r, echo = FALSE}
ExMin = brfss %>% filter(!is.na(BMI_in),!is.na(exract11), !is.na(X_paindx1),!is.na(padur1_),
                         !is.na(pa1min_), !is.na(X_totinda), !is.na(X_pa150r2), !is.na(X_pa300r2), !is.na(X_rfbmi5)) %>% 
    dplyr::select(BMI_in,exract11,X_paindx1,padur1_, pa1min_, X_totinda, X_pa150r2, X_pa300r2, X_rfbmi5)

ggplot(ExMin, aes(padur1_,BMI_in , color=X_rfbmi5))+
    geom_point(aes(color = X_rfbmi5), alpha = 1/4)+
    geom_jitter(aes(color = X_rfbmi5), alpha = 1/3)+
    coord_cartesian(xlim = c(0,200))+
    
    theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(hjust = 1),
        aspect.ratio = 3 / 5 
        ) +
  labs(x = "Total Weekly exercise in Minutes", y = "BMI", color= "Obesity")+
    ggtitle("Most do not meet CDC's exercise time recomedations")
```

##
```{r, echo = FALSE}
HlthEmploy = brfss %>% 
   filter(!is.na(BMI_in), !is.na(poorhlth), !is.na(physhlth),!is.na(menthlth), !is.na(sex), !is.na(employ1), !is.na(genhlth)) %>% 
    dplyr::select(BMI_in, poorhlth,physhlth, menthlth, X_age80, sex , employ1, genhlth) %>% 
    mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))

ggplot(HlthEmploy, aes(BMI_in, poorhlth))+
    geom_point(aes(color = employ1), alpha = 1/2)+
    facet_wrap(~genhlth, ncol = 1)+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 90, hjust = 1)
        )+
    labs(x = "BMI", y = "# Days in poor health", color = "Employment")
```

#Chronic Diseases

##

```{r,echo=FALSE}


diseases <- c('Diabetes','Asthma','High Blood Pressure','Cancer','High Cholesterol','COPD','Depression','Hearth Attack','Stroke')
dpercent <- c(.116,.124,.372,.09,.347,.077,.18,.055,.038)

alldisease <- data.frame(diseases, dpercent)%>%mutate(diseases = ordered(factor(diseases, levels = c("Stroke", "Hearth Attack", "COPD", "Cancer",  "Diabetes","Asthma","Depression","High Cholesterol", "High Blood Pressure")))) 



ggplot(data = alldisease, aes(x = factor(diseases), y = dpercent, fill=diseases)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) + 
 geom_text(aes(label = scales::percent(round((dpercent), 2)), y = dpercent - .02 ), hjust = 0.3, vjust = -2, size = 3.3)+
labs(x="General Health",y="Percentage")+
  
 ggtitle("The land of the free, the land of the disease")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
        legend.position="none",
       axis.text.x = element_text(angle = 45, hjust = 1)
       )
```

##

```{r,echo=FALSE}
dat.temp2 <- brfss %>% select(c(bphigh4, X_rfbing5, genhlth)) %>% filter(X_rfbing5!='NA',bphigh4!='NA',genhlth!='NA')%>%mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))%>% filter(bphigh4=='Yes' | bphigh4=='No')
total.grp <- dat.temp2 %>% group_by(genhlth) %>% count(genhlth)

dat.plot <- dat.temp2 %>% group_by(genhlth, genhlth, bphigh4) %>% count(genhlth)
dat.plot <- merge(dat.plot, total.grp, by = "genhlth")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(genhlth), y = Percentage,  fill = bphigh4)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), position=position_dodge(width = 0.9),vjust= -1,size=4)+
labs(x="General Health",fill="High Blood Pressure")+

ggtitle("Feel the pressure yet?")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
```

##

```{r,echo=FALSE}
dat.temp2 <- brfss %>% filter(X_rfchol=='Yes' | X_rfchol=='No') %>% select(c(X_rfchol, X_rfbing5, genhlth)) %>% filter(X_rfbing5!='NA',X_rfchol!='NA',genhlth!='NA')%>%mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))
total.grp <- dat.temp2 %>% group_by(genhlth) %>% count(genhlth)

dat.plot <- dat.temp2 %>% group_by(genhlth, genhlth, X_rfchol) %>% count(genhlth)
dat.plot <- merge(dat.plot, total.grp, by = "genhlth")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(genhlth), y = Percentage,  fill = X_rfchol)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ),position=position_dodge(width = 0.9),vjust= -1,size=4)+
labs(x="General Health",fill="High Cholesterol")+

ggtitle("High cholesterol respondents report good or better health?")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
```



##

```{r,echo=FALSE}
dat.temp2 <- brfss %>% filter(diabete3=='Yes' | diabete3=='No') %>% select(c(addepev2, X_rfbing5, genhlth)) %>% filter(X_rfbing5!='NA',addepev2!='NA',genhlth!='NA')%>%mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))
total.grp <- dat.temp2 %>% group_by(genhlth) %>% count(genhlth)

dat.plot <- dat.temp2 %>% group_by(genhlth, genhlth, addepev2) %>% count(genhlth)
dat.plot <- merge(dat.plot, total.grp, by = "genhlth")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(genhlth), y = Percentage,  fill = addepev2)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ),position=position_dodge(width = 0.9),vjust= -1,size=4)+
labs(x="General Health",fill="Depresion")+

ggtitle("Happy is healthy")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
```




##

```{r,echo=FALSE}
dat.temp2 <- brfss   %>% filter(diabete3=='Yes' | diabete3=='No') %>% select(c(diabete3, X_rfbing5, genhlth)) %>% filter(X_rfbing5!='NA',diabete3!='NA',genhlth!='NA')%>%mutate(genhlth = ordered(factor(genhlth, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))))
total.grp <- dat.temp2 %>% group_by(genhlth) %>% count(genhlth)

dat.plot <- dat.temp2 %>% group_by(genhlth, genhlth, diabete3) %>% count(genhlth)
dat.plot <- merge(dat.plot, total.grp, by = "genhlth")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(genhlth), y = Percentage,  fill = diabete3)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), position=position_dodge(width = 0.9),vjust= -1,size=4)+
labs(x="General Health",fill="Diabetes")+

ggtitle("I tought I was healthy")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
```

##

```{r,echo=FALSE}

Regions2 <- c('Midwest Region','Northeast Region','South Region','West Region')
rpercent <- c(.269,.154,.397,.18)

allregion <- data.frame(Regions2, rpercent)#%>%mutate(diseases = ordered(factor(diseases, levels = c("Stroke", "Hearth Attack", "COPD", "Cancer",  "Diabetes","Asthma","Depression","High Cholesterol", "High Blood Pressure")))) 



ggplot(data = allregion, aes(x = factor(Regions2), y = rpercent, fill=Regions2)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) + 
 
  
 geom_text(aes(label = scales::percent(round((rpercent), 2)), y = rpercent - .02 ),position=position_dodge(width = 0.9),vjust= -1.5,size=4)+
labs(x="General Health",y="Percentage")+
 ggtitle("Where is the highest diabetes prevalence?")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
        legend.position="none",
       axis.text.x = element_text(angle = 45, hjust = 1)
       )

```


##

```{r,echo=FALSE}
dat.temp <- brfss %>% filter(diabete3=='Yes' | diabete3=='No') %>% select(c(sex, X_rfbing5, diabete3)) %>% filter(X_rfbing5!='NA')

total.grp <- dat.temp %>% group_by(X_rfbing5) %>% count(X_rfbing5)

dat.plot <- dat.temp %>% group_by(X_rfbing5, diabete3, sex) %>% count(diabete3)
dat.plot <- merge(dat.plot, total.grp, by = "X_rfbing5")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(diabete3), y = Percentage,  fill = sex)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 facet_grid(.~ X_rfbing5) + 
  
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage - .02 ), position=position_dodge(width = 0.9),vjust= -1,size=4)+
labs(x="Diabetes")+

ggtitle("Fit women at higher risk of dibetes, for men it is obesity")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
  
  
```


##

```{r,echo=FALSE}
ageinsul2 <- brfss %>% select(c(X_age80,conv_weight,sex,diabete3)) %>% filter(X_age80!='NA',conv_weight!='NA',diabete3!='NA')%>% filter(diabete3=='Yes' | diabete3=='No')
ggplot(ageinsul2, aes(X_age80,conv_weight)) +
  geom_jitter(aes(color=diabete3),alpha=.9)+
  coord_cartesian(xlim = c(19,80))+
      ggtitle("Risk of diabetes increases at 40")+
  labs(x="Age",y= "Weight (Pounds)" ,color="Diabetes")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )

```



##

```{r,echo=FALSE}
ageinsul <- brfss %>% select(c(insulin,diabage2,X_age80)) %>% filter(insulin!='NA',diabage2!='NA', X_age80!='NA')
ggplot(ageinsul, aes(X_age80,diabage2))+
  geom_point(aes(color=insulin))+
  coord_cartesian(xlim = c(19,75))+
  labs(x= "Current Age", y="Age when Diagnosed",color= "Insulin Intake")+
    ggtitle("The older you are when diagnosed, more likely not to take insulin")+
   theme(panel.background = element_blank(),
       axis.line = element_line(),
       axis.text.x = element_text()
       )
  
```



# Behavioral Risks

##
<img src = "C:/Users/Biljana/Desktop/got_advice.jpeg" alt = "Image not loaded",width = "768", height = "600" />

##
```{r, echo = FALSE}
brfss$income2_factor = factor(brfss$income2, 
                       levels = c("Less than $10,000", "Less than $15,000", "Less than $20,000", "Less than $25,000" ,
                                  "Less than $35,000", "Less than $50,000", "Less than $75,000", "$75,000 or more"))


Diabetes_income = brfss %>% dplyr::select(income2_factor, diabete3) %>% filter(diabete3 == "Yes" | diabete3 == "No")

Diabetes_income_clean = Diabetes_income[complete.cases(Diabetes_income), ]

total.grp <- Diabetes_income_clean %>% group_by(income2_factor) %>% count(income2_factor)

dat.plot <- Diabetes_income_clean %>% group_by(income2_factor, diabete3) %>% count(income2_factor)
dat.plot <- merge(dat.plot, total.grp, by = "income2_factor")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(income2_factor), y = Percentage,  fill = diabete3)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
 
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage + .009 ), position=position_dodge(width=0.9), vjust=-0.25, size=3.1)+
   
  labs( x="Income Level $", title = "More money, Less problems!", fill = "Diabetes") + 
  scale_x_discrete( labels = c( "Less than $10,000" = "<10K" , "Less than $15,000" = "10K-15K" , 
                                "Less than $20,000" = "15K-20K" , "Less than $25,000" = "20K-25K",  
                                "Less than $35,000" = "25K-35K", "Less than $50,000" = "35K-50K",
                                "Less than $75,000" = "50K-75K", "$75,000 or more" = "Over 75K" )) +
  theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        )


```


##
```{r, echo = FALSE}
brfss$income2_factor = factor(brfss$income2, 
                       levels = c("Less than $10,000", "Less than $15,000", "Less than $20,000", "Less than $25,000" ,
                                  "Less than $35,000", "Less than $50,000", "Less than $75,000", "$75,000 or more"))


DEPRESSION_income = brfss %>% dplyr::select(income2_factor, addepev2)

DEPRESSION_income_clean = DEPRESSION_income[complete.cases(DEPRESSION_income), ]

total.grp <- DEPRESSION_income_clean %>% group_by(income2_factor) %>% count(income2_factor)

dat.plot <- DEPRESSION_income_clean %>% group_by(income2_factor, addepev2) %>% count(income2_factor)
dat.plot <- merge(dat.plot, total.grp, by = "income2_factor")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(income2_factor), y = Percentage,  fill = addepev2)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage + .009 ), position=position_dodge(width=0.9), vjust=-0.25, size=3.1)+
  
  labs( x="Income Level $", title = "With poverty comes depression", fill = "Depression") + 
  scale_x_discrete( labels = c( "Less than $10,000" = "<10K" , "Less than $15,000" = "10K-15K" , 
                                "Less than $20,000" = "15K-20K" , "Less than $25,000" = "20K-25K",  
                                "Less than $35,000" = "25K-35K", "Less than $50,000" = "35K-50K",
                                "Less than $75,000" = "50K-75K", "$75,000 or more" = "Over 75K" ))+
  theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        )
```

##

```{r, echo=FALSE}
brfss$income2_factor = factor(brfss$income2, 
                       levels = c("Less than $10,000", "Less than $15,000", "Less than $20,000", "Less than $25,000" ,
                                  "Less than $35,000", "Less than $50,000", "Less than $75,000", "$75,000 or more"))


CANCER_income = brfss %>% dplyr::select(income2_factor, chcocncr)

CANCER_income_clean = CANCER_income[complete.cases(CANCER_income), ]

total.grp <- CANCER_income_clean %>% group_by(income2_factor) %>% count(income2_factor)

dat.plot <- CANCER_income_clean %>% group_by(income2_factor, chcocncr) %>% count(income2_factor)
dat.plot <- merge(dat.plot, total.grp, by = "income2_factor")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = factor(income2_factor), y = Percentage,  fill = chcocncr)) + geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
  
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage + .009 ), position=position_dodge(width=0.9), vjust=-0.25, size=3.1)+
  
  labs( x="Income Level $", title = "Money won't save you (from cancer) ", fill = "Cancer") + 
  scale_x_discrete( labels = c( "Less than $10,000" = "<10K" , "Less than $15,000" = "10K-15K" , 
                                "Less than $20,000" = "15K-20K" , "Less than $25,000" = "20K-25K",  
                                "Less than $35,000" = "25K-35K", "Less than $50,000" = "35K-50K",
                                "Less than $75,000" = "50K-75K", "$75,000 or more" = "Over 75K" ))+
  theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        )
```

##

```{r, echo = FALSE}
brfss$income2_factor = factor(brfss$income2, 
                       levels = c("Less than $10,000", "Less than $15,000", "Less than $20,000", "Less than $25,000" ,
                                  "Less than $35,000", "Less than $50,000", "Less than $75,000", "$75,000 or more"))

health_income = brfss %>%       
  dplyr::select(income2_factor, poorhlth) 

health_income_clean = health_income[complete.cases(health_income), ]
ggplot(health_income_clean, aes(x = income2_factor, y = poorhlth)) + 
  
  geom_boxplot(outlier.shape = NA) + 
  geom_point(stat = "summary", fun.y = "mean", color = "blue", size = 4)+
  labs(x = "Income Level", y = "Number of days of bad health", title = " Get rich or die tryin'")+
   theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        ) +
  scale_x_discrete( labels = c( "Less than $10,000" = "<10K" , "Less than $15,000" = "10K-15K" , 
                                "Less than $20,000" = "15K-20K" , "Less than $25,000" = "20K-25K",  
                                "Less than $35,000" = "25K-35K", "Less than $50,000" = "35K-50K",
                                "Less than $75,000" = "50K-75K", "$75,000 or more" = "Over 75K" ))
```

##
<img src = "C:/Users/Biljana/Desktop/smoker.jpeg" alt = "Image not loaded", width = "768", height = "600" />

##
```{r, echo = FALSE}
CANCER_smokers = brfss %>% dplyr::select(X_smoker3, chcocncr)

CANCER_smokers_clean = CANCER_smokers[complete.cases(CANCER_smokers), ]

total.grp <- CANCER_smokers_clean %>% group_by(X_smoker3) %>% count(X_smoker3)

dat.plot <- CANCER_smokers_clean %>% group_by(X_smoker3, chcocncr) %>% count(X_smoker3)
dat.plot <- merge(dat.plot, total.grp, by = "X_smoker3")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = X_smoker3, y = Percentage,  fill = chcocncr)) + 
  
  geom_bar(stat = "identity", position = "dodge") +
  
 scale_y_continuous(labels=percent) +
 geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage + .009 ), position=position_dodge(width=0.9), vjust=-0.25, size=3.1)+
  
  labs(x = NULL, title = "Smoking data is problematic", fill = "Cancer") + 
  scale_x_discrete( labels = c( "Current smoker - now smokes some days" = "Occasional Smoker", 
                                 "Current smoker - now smokes every day" = "Smoker" )) +
  
  theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        )
```

##
```{r, echo = FALSE}
brfss$bmi_factor = factor(brfss$X_bmi5cat,
                       levels = c( "Obese" , "Overweight" , "Normal weight" , "Underweight" ))


Diabetes_WEIGHT = brfss %>% dplyr::select(bmi_factor, diabete3) %>% filter(diabete3 == "Yes" | diabete3 == "No")

Diabetes_WEIGHT_clean = Diabetes_WEIGHT[complete.cases(Diabetes_WEIGHT), ]

total.grp <- Diabetes_WEIGHT_clean %>% group_by(bmi_factor) %>% count(bmi_factor)

dat.plot <- Diabetes_WEIGHT_clean %>% group_by(bmi_factor , diabete3) %>% count(bmi_factor)
dat.plot <- merge(dat.plot, total.grp, by = "bmi_factor")
dat.plot <- mutate(dat.plot, Percentage = n.x/n.y)

ggplot(data = dat.plot, aes(x = bmi_factor, y = Percentage,  fill = diabete3)) + 
  
  geom_bar(stat="identity", position = "dodge") +
 scale_y_continuous(labels=percent) +
 
 
  geom_text(aes(label = scales::percent(round((Percentage), 2)), y = Percentage + .009 ), position=position_dodge(width=0.9), vjust=-0.25, size=3.1) +
   
  labs( x= NULL, title = "Gaining weight is risky business", fill = "Diabetes") +
  
  theme(panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(angle = 0, hjust = .5)
        )

```


   